---
title: "get-stress-genes"
output: html_document
---
Rmd to get list of stress response genes from the transcriptome v 3.1 with `blast` output and uniprot-SP-Go annotation. 

```{r}
library(dplyr)
library(tidyverse)
```

Read in the uniprot-SP-GO.sorted file:
```{r}
uniprot_SP_GO <- read.delim("http://owl.fish.washington.edu/halfshell/bu-alanine-wd/17-07-20/uniprot-SP-GO.sorted", sep = '\t', header = FALSE)
head(uniprot_SP_GO)
```

Rename columns:
```{r}
colnames(uniprot_SP_GO) <- c("uniprot_acc_ID", "V2", "V3", "V4", "V5", "V6", "V7", "V8", "V9", "V10", "V11", "V12")
head(uniprot_SP_GO)
```

Read in `blast` output from transcriptome v 3.1:
```{r}
blast <- read.delim("../analyses/BLAST-to-GOslim/_blast-sep.tab", header = FALSE)
head(blast)
```

Rename first and third column for `join`-ing purposes later on. 
```{r}
colnames(blast) <- c("Trinity_ID", "V2", "uniprot_acc_ID", "V4", "V5", "V6", "V7", "V8", "V9", "V10", "V11", "V12", "V13", "V14")
head(blast)
```

`join` the `blast` output with the uniprot-SP-GO by uniprot_acc_ID:
```{r}
blastgo <- left_join(blast, uniprot_SP_GO, by = "uniprot_acc_ID")
head(blastgo)
```
50038 rows with 25 columns

Read in the table from the BLAST to GOslim jupyter notebook called "Blastquery-GOslim.tab":
```{r}
blastquery <- read.delim("../analyses/BLAST-to-GOslim/Blastquery-GOslim.tab", sep = '\t', header = FALSE)
head(blastquery)
```
459411 rows.

Add column names:
```{r}
colnames(blastquery) <- c("Trinity_ID", "GO_ID", "GOslim", "biological_process")
head(blastquery)
```

Subset the GOslim terms that have the word "stress" in them. Only the term "stress response" has the term "stress":
```{r}
stress <- blastquery %>%
  filter(str_detect(GOslim, "stress"))
head(stress)
```
11046 Trinity IDs.

`join` the stress list with the annotated `blast` output with uniprot (blastgo):
```{r}
stress_annot <- left_join(stress, blastgo, by = "Trinity_ID")
head(stress_annot)
```
36735 rows. 28 columns.       
Some Trinity IDs have multiple GO IDs associated with it - hence the extra rows. 

write out the annotated stress response table:
```{r}
#write.table(stress_annot, "../analyses/BLAST-to-GOslim/transcriptome-stress-genes.tab", sep = '\t', quote = FALSE, row.names = FALSE)
```
Wrote out 08/20/2020. Commented out code. 









