---
title: "GOslim-pie"
output: html_document
---
Rmd to create a pie chart of GOslim terms from transcriptome v 3.1. 

## Read in file of counts for pie chart:
```{r}
gopie <- read.delim("../analyses/BLAST-to-GOslim/GOslim-P-pie.txt", header = FALSE)
gopie
```

Will not include terms: "other biological processes" or "other metabolic processes" 

pie chart colors:
```{r}
piecolors <- RColorBrewer::brewer.pal(12, "Set3")
```

Make the pie chart with percentages:
https://www.statmethods.net/graphs/pie.html 
```{r}
slices <- c(8841, 6507, 3858, 26964, 23297, 23265, 2129, 1768, 17143, 15006, 10910, 10319) #create vector of counts of GOslims based on the file read in above; don't use numbers associated with "other biological processes" or "other metabolic processes"
lbls <- c("cell cycle and proliferation", "DNA metabolism", "death", "developmental processes", "RNA metabolism", "cell organization and biogenesis", "cell-cell signaling", "cell adhesion", "protein metabolism", "transport", "stress response", "signal transduction") #create vector of GOslim terms in the same order as the number of counts; don't include "other biological processes" or "other metabolic processes"
pct <- round(slices/sum(slices)*100) #get percentages of each GOslim term
lbls <- paste(lbls, pct) #add percents to labels
lbls <- paste(lbls, "%", sep = "") #ad % to labels
pie(slices,labels = lbls, col = piecolors)
```

Save pie chart as a pdf:
```{r}
pdf("../analyses/BLAST-to-GOslim/GOslim-pie-transcriptome-v3.1.pdf")
slices <- c(8841, 6507, 3858, 26964, 23297, 23265, 2129, 1768, 17143, 15006, 10910, 10319) #create vector of counts of GOslims based on the file read in above; don't use numbers associated with "other biological processes" or "other metabolic processes"
lbls <- c("cell cycle and proliferation", "DNA metabolism", "death", "developmental processes", "RNA metabolism", "cell organization and biogenesis", "cell-cell signaling", "cell adhesion", "protein metabolism", "transport", "stress response", "signal transduction") #create vector of GOslim terms in the same order as the number of counts; don't include "other biological processes" or "other metabolic processes"
pct <- round(slices/sum(slices)*100) #get percentages of each GOslim term
lbls <- paste(lbls, pct) #add percents to labels
lbls <- paste(lbls, "%", sep = "") #ad % to labels
pie(slices,labels = lbls, col = piecolors)
```

----

Make pie including the "other..." GOslim terms to get more accurate percentages:

From here: https://stackoverflow.com/questions/33259275/how-to-place-the-labels-further-from-pie-chart


New pie colors for 14 terms:
```{r}
newpiecolors <- rainbow(14)
```

```{r}
slices2 <- c(8841, 6507, 3858, 27904, 26964, 23297, 23265, 2129, 20834, 1768, 17143, 15006, 10910, 10319) #create vector of counts of GOslims based on the file read in above
lbls2 <- c("cell cycle and proliferation", "DNA metabolism", "death", "other biological processes", "developmental processes", "RNA metabolism", "cell organization and biogenesis", "cell-cell signaling", "other metabolic processes", "cell adhesion", "protein metabolism", "transport", "stress response", "signal transduction") #create vector of GOslim terms in the same order as the number of counts
pct2 <- round(slices2/sum(slices2)*100) #get percentages of each GOslim term
lbls2 <- paste(lbls2, pct2) #add percents to labels
lbls2 <- paste(lbls2, "%", sep = "") #ad % to labels
pie(slices2, labels = lbls2, col = newpiecolors, radius = 0.7)
```

Saved heatmap by right clicking > Save Image As, download as png, then open in preview, export > PDF, save. 

```{r}
#pdf("../analyses/BLAST-to-GOslim/allterms.GOslim-pie-transcriptome-v3.1.pdf", width = 11, height = 8.5)
#slices2 <- c(8841, 6507, 3858, 27904, 26964, 23297, 23265, 2129, 20834, 1768, 17143, 15006, 10910, 10319) #create vector of counts of GOslims based on the file read in above
#lbls2 <- c("cell cycle and proliferation", "DNA metabolism", "death", "other biological processes", "developmental processes", "RNA metabolism", "cell organization and biogenesis", "cell-cell signaling", "other metabolic processes", "cell adhesion", "protein metabolism", "transport", "stress response", "signal transduction") #create vector of GOslim terms in the same order as the number of counts
#pct2 <- round(slices2/sum(slices2)*100) #get percentages of each GOslim term
#lbls2 <- paste(lbls2, pct2) #add percents to labels
#lbls2 <- paste(lbls2, "%", sep = "") #ad % to labels
#pie(slices2, labels = lbls2, col = newpiecolors, radius = 0.6)
```










