---
title: "fig2 and fig3 cluster join"
output: html_document
---
Rmd to `join` the cluster heatmap information from Fig 2 in the manuscript (heatmap of the 123 DEGs in the 4 pooled samples) with the cluster heatmap information of the individual crab sequenced over time (Fig 3.) Only want the 94 that match.


```{r}
library(dplyr)
library(tidyverse)
```

Read in annotated temp-influenced infection DEG table (Fig 2):
```{r}
temp.annot <- read.delim("../analyses/pheatmap/contrasttemp_DEGs-clusters-annot.tab")
head(temp.annot)
```

Read in single crab gene cluster annotated list:
Read in the table of genes with clusters and annotation from the individual crab over time heatmap (from Rmd: heatmaps-single_crab-over-time.Rmd) 
```{r}
singlecrab.genes <- read.delim("../analyses/pheatmap/single_crab-clusters-blast-GO.tab")
head(singlecrab.genes)
```

13954 rows. 

Filter out the Trinity_ID column and all others except the count column, because that will be redundant in the final `join`-ed table. 
```{r}
singlecrab.genes <- select(singlecrab.genes, "Trinity_ID", "cluster")
head(singlecrab.genes)
```

Rename cluster column and corrent the typo of "Geme.names", and add '.Fig3" to end of "singlecrab.gene.cluster" column name to note the manuscript figure number:
```{r}
colnames(singlecrab.genes) <- c("Trinity_ID", "singlecrab.gene.cluster.Fig3")
head(singlecrab.genes)
```

`join` the two tables by "Trinity_ID"
```{r}
temp.degs.single <- left_join(temp.annot, singlecrab.genes, by = "Trinity_ID")
head(temp.degs.single)
```

Rearrange so the two cluster columns are next to each other. 

```{r}
temp.degs.single <- temp.degs.single %>% select("Trinity_ID", "X380822", "X380823", "X380824", "X380825", "cluster", "singlecrab.gene.cluster.Fig3", "swissprot", "uniprot_acc_ID", "gene_id", "pident", "length", "mismatch", "gapopen", "qstart", "qend", "sstart", "send", "evalue", "bitscore", "Entry.name", "Status", "Protein.names", "Gene.names", "Organism", "Length", "Gene.ontology.biological.process", "Gene.ontology.cellular.component", "Gene.ontology.GO", "Gene.ontology.molecular.function", "Gene.ontology.IDs")
head(temp.degs.single)
```

Rename column name column for the temperature contrast deg clusters (Fig. 2):
```{r}
colnames(temp.degs.single) <- c("Trinity_ID", "380822", "380823", "380824", "380825", "temp.contrast.cluster.Fig2", "singlecrab.gene.cluster.Fig3", "swissprot", "uniprot_acc_ID", "gene_id", "pident", "length", "mismatch", "gapopen", "qstart", "qend", "sstart", "send", "evalue", "bitscore", "Entry.name", "Status", "Protein.names", "Gene.names", "Organism", "Length", "Gene.ontology.biological.process", "Gene.ontology.cellular.component", "Gene.ontology.GO", "Gene.ontology.molecular.function", "Gene.ontology.IDs")
head(temp.degs.single)
```

Write out table to analyses:
```{r}
#write.table(temp.degs.single, "../analyses/pheatmap/temp-influenced-infectionDEGs-singlecrab-overtime.tab", sep = '\t', quote = FALSE, row.names = FALSE)
```
Wrote out 02/11/2021. Commented out code. 


